\section{Program transformation}
\label{sec-transformation}

In this section we present a program transformation that realizes the
function \monitor presented in Section~\ref{sec-semantics}.
To implement the state passing between judgemets we apply a so-called
logger monad.

\input{sec_logger}

\paragraph{\bf Transformation of types}

We transform each program expression into a monadic expression that
keeps track of the state that results in the judgement augmentation.
Figure~\ref{fig-types-prods} presents the function \monadic\ that maps
types of expressions in the original program to types of the
transformed program.
Function \monadic\ indicates that a tranformed program is a \miniocaml\
function that takes an initial state and returns a \mbox{pre-monadic}
program value together with a final state.

\begin{figure}[t]
  \linespread{1}
  \begin{minipage}[t]{\columnwidth}
    \small
    \begin{tabular}{l@{\ }l}
      1 & \algLet \algRec \monadic\ \type\ = (\premonadic\ \type) \moncstr \\[\jot]
      & \\[\jot]
      2 & \algAnd \premonadic\ = \algFunction \\[\jot]
      3 & \tabT $|$ \ttt{\type[1] -> \type[2]}\
      \algArrow \ttt{$(\premonadic\ \type[1])$ -> $(\premonadic\
        \type[2])$ m} \\[\jot]
      % 6 & \tabT $|$ \ttt{\type[1]\ *\ \ldots\ *\ \type[n]}\
      % \algArrow \ttt{$(\premonadic\ \type[1])$\ *\ \ldots\ *\
      %   $(\premonadic\ \type[n])$} \\[\jot] 
      4 & \tabT $|$ \ttt{(\type[1],\ \ldots,\ \type[n])\ \tycstr}\ 
      \algArrow \ttt{($(\premonadic\ \type[1])$,\ \ldots,\ $(\premonadic\
      \type[n])$)\ \tycstr} \\[\jot]
      5 & \tabT $|$ \type\ \algArrow \type
    \end{tabular}
  \end{minipage}
  \caption{Type transformation function \monadic.}
  \label{fig-types-prods}
\end{figure}

For example, consider the following applications of \monadic.
%
\begin{align*}
  \monadic\ \texttt{(int -> (int -> int))} = \;
  \begin{array}[t]{@{}l@{}}
    \texttt{(int -> ((\premonadic\ int -> int) m)) m}\\[\jot]
    \texttt{(int -> ((int -> int m) m)) m}
    % (int -> ((prem int -> int) m)) m
    % (int -> ((int -> int m) m)) m
  \end{array}
\end{align*}
%
\input{fig_tsel}

\input{fig_prod_constr}

\paragraph{\bf Transformation of expressions}


We present the transformation function \product\ in
Figure~\ref{fig-prod-alg}. 
\product uses two auxiliary functions $\tselEnter$ and $\tselExit$
shown in Figure~\ref{fig-tsel}.

For a expression \expr, \product\ traverses the abstract syntax tree
of \expr\ and gives a core monadic expression that evaluates the user
program together with two state transform operations.
\product generates \miniocaml expressions using the $\quo{\cdot}$
function.
For example, $\quo{\ttt{let x = 1 in 1}}$ emits the expression
\ttt{let x = 1 in 1}.
Within $\quo{\cdot}$ we can perform an evaluation by
applying~$\aq{\cdot}$.
For example, $\quo{\ttt{let x = \aq{\quo{1+2}} in 1}}$ emits $\ttt{let
x = 1+2 in 1}$.

The important case is the tranformation of
$\theFunction/\theArity$-applications.
Such applications are recognized in $\tselEnter$ and $\tselExit$.
The emitted code either saves the argument values into the state, or
propagates further the current state.
Furthermore, $\tselEnter$ performs a check if the snapshot stored in a
state together with the arguments of a
$\theFunction/\theArity$-application satisfy the transition invariant
candidate~$\theTI$.
This check is guarded by the condition that the snapshot must have
been stored previously.

We show an example application of \product in Figure~\ref{fig-ex-prod}
for analysing $\texttt{f}/1$-applications.
First, we present subexpressions of the program and then show the
result of the application of \product on them (we have partially
simplified the transformed expressions to improve readability).
%
\begin{figure}[!t]
  \centering
{
\small
\begin{verbatim}
e = let rec f x = if x > 0 then f (x - 1) else fun y -> f x y in f 1
e1 = if x > 0 then f (x - 1) else fun y -> f x y
e2 = x > 0
e3 = 0
e4 = (>) x
e5 = x
e6 = (>)
e7 = f (x - 1)
...

Transform e = let rec f = fun x -> 'Transform e1' in 'Transform e20'
Transform e1 = 'Transform e2' >>= fun x2 -> (if x2 then 'Transform e7'
                                             else 'Transform e14')
Transform e2 = fun s_full -> ('Transform e3' >>= fun x3 ->
                              'Transform e4' >>= fun x4 ->
                              update (fun s -> s) >>= fun () s_partial ->
                             (x4 x3 >>= fun xapp2 ->
                              update (fun s -> s) >>= fun () ->
                              unit xapp2) s_partial) s_full
Transform e3 = unit 0
Transform e4 = fun s_full -> ('Transform e5' >>= fun x5 ->
                              'Transform e6' >>= fun x6 ->
                              update (fun s -> s) >>= fun () s_partial ->
                             (x6 x5 >>= fun xapp4 ->
                              update (fun s -> s) >>= fun () ->
                              unit xapp4) s_partial) s_full
Transform e5 = unit x
Transform e6 = unit (fun z1 -> unit (fun z2 -> unit (z1 > z2)))
Transform e7 = fun s_full -> ('Transform e8' >>= fun x8 ->
                              'Transform e13' >>= fun x13 ->
                              update (fun ( _, m ) ->
                                let m_c, m_1 = m in
                                if m_c then assert ( x8 > 0 && m_1 > x8 );
                                x8, if not m_c && nondet () then true, x8 else m
                              ) >>= fun () s_partial ->
                             (x13 x8 >>= fun xapp7 ->
                              update (fun _ -> s_full) >>= fun () ->
                              unit xapp7) s_partial) s_full
...
\end{verbatim}
}
  \caption{Example application of \product\ with $N = 1$ and \mbox{$\theTI = (\
      a_1 > 0 \wedge m_1 > a_1 \ )$}. Given a stored snapshot, the
    transformed application \ttt{f (x - 1)} checks that the current
    actual satisfies $\theTI$.}
  \label{fig-ex-prod}
\end{figure}
%

We establish a relationship between augmented evaluation trees and
evaluation trees of the transformed program in the following theorem.
%
\begin{theorem}[\product implements \monitor]
\label{thm-product}
%
A pair $(v_1, \dots, v_\theArity)$ and~$(u_1, \dots, u_\theArity)$ is
obtained from the augmented evaluation tree as described in
Theorem~\ref{thm-monitor}
if and only if a judgement of the following form appears in the
evaluation tree of the
program obtained by applying \product:
%
\begin{center}
  \mevj{\evj{\ectx}{$\theFunction$\ \expr[1]\ \_\!\_\ \dots\
      \expr[\theArity]\ s}{\val}}{\stupa}{}
\end{center}
%
such that $\tsf{eval}\ \ectx\ s = (true, (v_1, \dots, v_\theArity))$
and for each $i \in 1..\theArity$ we have \mbox{$\tsf{eval}\ \ectx\ \expr[i]
= u_i$}.
%
\end{theorem}
%

The following corollary of Theorem~\ref{thm-product} allows one to
rely on the assertion validity in the transformed program to implement
the binary reachability analysis of the original program.
%
\begin{theorem}[Binary reachability analysis as assertion checking]
\label{thm-assert}
%
Each pair $(v_1, \dots, v_\theArity)$ and~$(u_1, \dots, u_\theArity)$
in the $\theFunction/\theArity$-recursion relation of the program 
satisfies $\theTI$ if and only if the assertion inserted by
$\tselEnter$ is valid in the transformed program.
%
\end{theorem}
%


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: 
