\iffalse

\newpage
\section{Benchmarks of \bfc}

We experimented on the benchmarks \ttt{cprover-PN} and
\ttt{cprover\_software\_analysis} available at \\
\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
Table~\ref{bfc-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{7cm} | r | r | r | r | } %
    \hline
    Method & pos & neg & don't know & TO \\
    \hline
    Safety (over real numbers or naturals)      & 23 &  0 & 54 &  0 \\
    Invariant \& Invariant with Minimization    & 23 &  0 & 54 &  0 \\
    Safety by Refinement                        & 23 &  0 & 54 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \bfc}
\label{bfc-experiments}
\end{table}

\section{Benchmarks of \mist}

We experimented on the benchmarks available at \\
\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
and \\
\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
Table~\ref{mist-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{6cm} | r | r | r | r | }
    \hline
    Method & pos & neg & don't know & TO \\
    \hline
    Safety (over real numbers or naturals)      & 14 &  0 & 13 &  0 \\
    Invariant \& Invariant with Minimization    & 14 &  0 & 13 &  0 \\
    Safety by Refinement                        & 20 &  0 &  7 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \mist}
\label{mist-experiments}
\end{table}

\section{Benchmarks of \iic}

We experimented on the benchmarks available at \\
\ttt{https://svn.rse.mpi-sws.org/RSE/IC3forWSTS/code/trunk/examples/soter}.
The set of benchmarks consists 38 benchmarks, 28 positive and 10 negative.
Table~\ref{iic-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{6cm} | r | r | r | r | }
    \hline
    Method & pos & neg & don't know & TO \\
    \hline
    Safety (over real numbers or naturals)      & 2 & 0 & 36 & 0 \\
    Invariant \& Invariant with Minimization    & 2 & 0 & 26 & 0 \\
    Safety by Refinement                        & 2 & 0 & 36 & 0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \iic}
\label{iic-experiments}
\end{table}

We constructed the following two invariants by applying the method
Invariant with Minimization.

\begin{itemize}
\item \verb?firewall__no_pred_called_with_zero__depth_1__fixed?: \verb?(assert (<= l739 0))?
\item \verb?firewall__no_pred_called_with_zero__depth_2__fixed?: \verb?(assert (<= l3849 0))?
\end{itemize}

\iic\ constructs the following two invariants.

\begin{itemize}
\item \verb=firewall__no_pred_called_with_zero__depth_1__fixed.spec=: \verb=(l739 < 1)=
\item \verb=firewall__no_pred_called_with_zero__depth_2__fixed.spec=: \verb=(l3849 < 1)=
\end{itemize}

\fi