\section{Preliminaries}
\label{sec_preliminaries}

A \emph{Petri net} is a tuple $(P, T, F, m_0)$, where $P$ is a set of
\emph{places}, $T$ is a (disjoint) set of \emph{transitions},
$F : (P\times T) \cup (T\times P) \to \set{0,1}$ is the \emph{flow function},
and $m_0: P \rightarrow \mathbb{N}$ is the initial marking.
%
For $x\in P\cup T$, the \emph{pre-set} is
$\pre{x}=\{y\in P\cup T\mid F(y,x) =1\}$
and the \emph{post-set} is $\post{x}=\{y\in P\cup T\mid F(x,y) = 1\}$.
We extend the pre- and post-set to a subset of $P \cup T$ as the union of
the pre- and post-sets of its elements.

A \emph{marking} of a Petri net is a function $m\colon P \to \mathbb{N}$,
which describes the number of tokens $m(p)$ in 
each place $p\in P$. 
Assuming an enumeration $p_1, \ldots, p_n$ of $P$, we
often identify $m$ and the vector $(m(p_1), \ldots, m(p_n))$.
For a subset $P'\subseteq P$ of places, we write $m(P') = \sum_{p\in P'} m(p)$.

A transition $t \in T$ is \emph{enabled at $m$} iff
for all $p \in \pre{t}$, we have $m(p) \ge F(p, t)$.
A transition $t$ enabled at $m$ may \emph{fire},
yielding a new marking $m'$ (denoted $m \xrightarrow{t} m'$),
where $m'(p) = m(p) + F(t,p) - F(p,t)$.

A sequence of transitions, $\sigma = t_1 t_2 \ldots t_r$ is an
$\emph{occurrence sequence}$ of $N$ iff there exist markings
$m_1, \ldots, m_r$ such that $m_0 \xrightarrow{t_1} m_1
\xrightarrow{t_2} m_2 \ldots \xrightarrow{t_r} m_r$. The marking
$m_r$ is said to be \emph{reachable} from $m_0$ by the occurrence
of $\sigma$ (denoted $m_0 \xrightarrow{\sigma} m_r$).

A \emph{property} $\varphi$ is a linear arithmetic constraint over the free variables $P$. 
The property $\varphi$ holds on a marking $m$ iff $m \models \varphi$.
% Examples of properties are $s_1 \le 2$, $s_1 + s_2 \ge 1$ and
% $((s_1 \le 1) \land (s_2 \ge 1)) \lor (s_3 \le 1)$.
% 
A Petri net $N$ satisfies a property $\varphi$ (denoted by $N \models \varphi$)
iff for all reachable markings $m_0 \xrightarrow{\sigma} m$, we have
$m \models \varphi$.
A property $\varphi$ is an \emph{invariant} of $N$ if it holds for every reachable marking.
A property is inductive if whenever $m\models \varphi$ and $m \xrightarrow{t} m'$ for some $t\in T$
and marking $m'$, we have $m'\models \varphi$.

% A trap is a set of places $S \subseteq P$ such that $\post{S} \subseteq \pre{S}$.
% If a trap $S$ is marked in $m_0$, i.e. $\sum_{p \in S} m_0(p) > 0$, then it is also marked in all reachable markings.

Petri nets are represented graphically as follows: places and transitions
are represented as circles and boxes, respectively. For $x, y \in P \cup T$,
there is an arc leading from $x$ to $y$ if{}f $F(x,y)=1$.

\input{fig_lamport_src}
\input{fig_lamport_net}

Petri nets can be used to model behavior of concurrent software. As an
example, consider Lamport's 1-bit
algorithm for mutual exclusion \cite{LamportJACM86}, shown in
Fig.~\ref{fig_lamport_src}. 
A Petri net model for the code is shown in Fig.~\ref{fig_lamport_net}.
The two grey blocks model the control flow of the two
processes. For instance, the token on place $p_1$ models the current
position of process 1 at program location $p_1$. The three places in the middle of the
diagram model the current values of the variables. For instance, a
token on place $notbit_1$ indicates that the variable ${\it bit}_1$ is currently
set to ${\it false}$.
The mutual exclusion property, which states that the two processes cannot be in the
critical section at the same time, corresponds to the property that
places $p_3$ and $q_5$ cannot both have a token at the same time.

